---
title: "Pollinator_outline"
author: "Michelle DePrenger-Levin"
date: "Monday, September 28, 2015"
output: html_document
---
```{r, echo=FALSE}
library(ggplot2)
```



DBG's data     
look at NMDS for distance to disturbance or ADD IN scales of disturbance within different circles around the populaiton     
lm(richness~contour lines*distance)     
at some threshold of surrounding disturbance does proximity to disturbance have an effect so a buffer should be put into effect at some threshold? 

```{r}
dbg_demo <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/2012_Black Hills/2012_data/Demographic data to Disturbance datacsv.csv"))

head(dbg_demo)
```


# DBG pollination 2012
```{r}
pol <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/2012_Black Hills/2012_data/2012_SCGL_Pollination_Study.csv"))

head(pol)

polsum <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/2012_Black Hills/2012_data/Pollination summary.csv"))

head(polsum)
```

Keep adding location for study points
```{r}
do.call(data.frame,lapply(split(pol, pol$Site), function(x){ 
  mean(rowSums(pol[,12:21], na.rm = TRUE))
}))

totpol <- cbind(apply(polsum[,-1],2,FUN=function(x) sum(x, na.rm = TRUE)))


#lm( ,data=pol)
```



2013 study data of ScGl size by distance group to a disturbance
```{r}
ind13 <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/2013_IndivData.csv"))

head(ind13)
```

I'm pretty sure that T.S means if the plant is touching another or is seperate
```{r}
site13 <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/2013_SiteData.csv"))

head(ind13)
```

```{r}
polcheck <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/2012_Black Hills/2012_data/2013-02-27_Pollinator_types_check.csv"))

head(polcheck)
```



```{r}
dist14 <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/2014_ScGl_disturbancedata.csv"))

head(dist14)
```


```{r}
germ14 <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/2014_ScGl_germinationHeightWidth.csv"))

head(germ14)
```

### 2014 germination of seeds in Becky's office    
Need to split the top rows that have date and temperatures from the bottom that give the number of seeds germinated per pot of near or far at each site

```{r}
germ14_all <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/Germination_ScGl_study2_2014.csv"))

head(germ14_all)
```


Access database "Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/Black Hills Disturbance.mdb"


```{r}
pol12 <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/Pollinators.csv"))

head(pol12)
```

From Access??
```{r}
samsize <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/qry_Sample Size_AllYears.csv"))

head(samsize)
```


### Soil samples from two north and two south sites 2015
```{r}
soil <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/2015_Black Hills/SoilSamplesResults.csv"))

head(soil)
```


##Seed germination near to far
```{r}
seed <- germ14_all[-c(1:3),]
names.seed <- germ14_all[3,]
names(seed) <- names.seed


```

Compare soil samples with seed germination rate, germination    
for ANOVA: fixed would be near far, random would be site since I want to generalize to other sites?   
Use Anova in <car> with type="III" which works best for unbalanced design
```{r}
maxgerm <- as.numeric(apply(seed[,10:103],1,function(x) max(x, na.rm=TRUE)))
seedmax <- data.frame(seed,maxgerm)
seed.soil <- merge(seedmax, soil, by.x = "X", by.y = "Site")

names(seed.soil)[2] <- "NF"
names(seed.soil)[1] <- "Site"

summary(lm(maxgerm~Site*NF, data=seed.soil))
```

##HSD Tukey test
```{r}
library(agricolae)
model <- lm(maxgerm~Site*NF, data=seed.soil)
HSD.test(model,list("NF","Site"), group=TRUE, console=TRUE)

```


Rate of germination?
```{r}
seed.soil[,5:103] <- lapply(seed.soil[,5:103], as.numeric)


lapply(split(seed.soil, list(seed.soil$Site,seed.soil$NF)), function(y){
  apply(y[,10:103],2,function(x)x/y$X.3)
})


germ.rate <- apply(seed.soil[,10:103],2,function(x)x/seed.soil$X.3)
head(germ.rate)
head(seed.soil)

ggplot()

seed2 <-read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BlackHillsDocuments/R code Black Hills/BlackHills_R_tables/Germination_20142015.csv"),na.strings="")

str(seed2)
days <- as.Date(substr(names(seed2[,-c(1:9)]),2,9), "%m.%d.%y")


names(seed2[,-c(1:9)]) <- days
names(seed2)


seed2[,5:103] <- lapply(seed2[,5:103], as.numeric)
str(seed2)

germ.rate <- apply(seed2[,10:103],2,function(x)x/seed2$NumSeeds)
head(germ.rate)
```




#Bio-Logic data     
bl_scgl has both glaucus and brevispinus but no wetlandicus
```{r}
bl_scgl <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BIO-Logic/scgl_2012.csv"), skip = 3, na.strings = "9999")

head(bl_scgl)
```

## Results 
Bio-logic    
Nbr: count of number of cactus within 10cm of marked plant    
Group: E(ast) or W(est) for scbr and N(orth) S(outh) for scgl     

#Calculate volume same way!
```{r}
bl_scgl$Vol <- pi*(bl_scgl$Max.Dia.^2)*bl_scgl$Max.Ht.

#ScGl and ScBr
ggplot(bl_scgl, aes(log(Vol+1), colour = Species, fill = Species, alpha = 0.5)) +
  geom_histogram()

summary(lm(log(Vol+1) ~ Meas.*Species, data = bl_scgl))
summary(lm(log(Vol+1) ~ Categ.*Species, data = bl_scgl))


biol <- bl_scgl[,c(1:21,46)]
table(biol$Species)


ggplot(biol, aes(Meas., log(Vol+1), colour = Species)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw() +
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  xlab("Distance from disturbance (m)") +
  ylab("log transformed volume")


```

Combind biologic and dbg   
```{r}
dbg_demo$Vol <- pi*(dbg_demo$Max.width^2)*dbg_demo$Max.Height
biol.2c <- biol[,c(1,5,22)]
dbg.2c <- data.frame(Species="SCGL", Meas.=dbg_demo[,12], Vol=dbg_demo$Vol)
comb_vol <- rbind(biol.2c, dbg.2c)

```

```{r}
ggplot(comb_vol, aes(Meas., log(Vol+1), colour = Species)) +
  geom_point() +
  stat_smooth(method = "lm") +
  theme_bw() +
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  xlab("Distance from disturbance (m)") +
  ylab("log transformed volume")


summary(lm(log(Vol+1) ~ Meas.*Species, data = comb_vol))
```


```{r}

head(dbg_demo)

ggplot(dbg_demo, aes(dist_disturb_1, log(Vol+1))) +
  geom_point(colour = "cyan3")+
  stat_smooth(method = "lm") +
  theme_bw() +
  theme(legend.justification=c(1,1),legend.position=c(1,1)) +
  xlab("Distance from disturbance (m)") +
  ylab("log transformed volume")
```

##Check same with ind13
```{r}
ind13$Vol <- pi*(ind13$Width^2)*ind13$Adj.Height.mm.
ind13_1 <- ind13[ind13$Vol != 0,]
density <- aggregate(ind13_1$Vol, 
                     list(Dist = ind13_1$DistanceGroup,
                          Site = ind13_1$Name,
                          Transect = ind13_1$Transect), length)

ggplot(density, aes(Dist,x))+
  geom_point()+
  stat_smooth(method="lm")+
  ylab(substitute(paste("Density of ", italic("Sclerocactus glaucus"))))+
  theme_bw()+
  facet_wrap(~Site)


ggplot(ind13_1, aes(DistanceGroup,log(Vol+1),colour=Name))+
  geom_point()+
  stat_smooth(aes(group=Name), method="lm")+
  facet_wrap(~Name)+
  theme_bw()


ggplot(ind13_1, aes(factor(DistanceGroup),log(Vol+1),colour=Name))+
  geom_boxplot()+
  facet_wrap(~Name)+
  theme_bw()


ggplot(ind13_1, aes(DistanceGroup,log(Vol+1)))+
  geom_point()+
  stat_smooth(method="lm")


ggplot(ind13_1, aes(factor(DistanceGroup),log(Vol+1), colour = Name))+
  geom_boxplot()

```

proportion of size in populations    
from DBG 2012-12-03 report: The average size of vegetative individuals at Black Hills sites was 54.6 cm3 (±10.2 SD). Black Hills individuals are significantly more likely to be reproductive once they reach 163.7 cm3 (± 2.8; t = -34.1, df = 1077.2, p < 0.001). Individuals are significantly smaller as the distance from a disturbance increases (r2 = 0.07, F(1,1158) = 82.2, p < 0.001; Figure 5a). 
```{r}
prop_sm <- aggregate(ind13_1$Vol,by=list(Site = ind13_1$Name,
                              Dist = ind13_1$DistanceGroup,
                              Transect = ind13_1$Transect), 
          function(x){
            length(x[x <16370])/length(x)
            })

ggplot(prop_sm, aes(factor(Dist),x))+
  geom_boxplot()+
  ylab("Proportion of young vegetative individuals")+
  xlab("Meters from a disturbance")

ggplot(prop_sm, aes(factor(Dist),x))+
  geom_boxplot()+
  stat_smooth(aes(group=Site), method="lm")+
  ylab("Proportion of young vegetative individuals")+
  xlab("Meters from a disturbance")+
  facet_wrap(~Site)+
  theme_bw()+
  theme(axis.text.x=element_text(angle=90,hjust=1))

summary(lm(x~Dist*Site, data=prop_sm))
```


## Need to get pollinator abundance and richness values from Biologic for analysis     
PBO: percent of bowl occupied out of 99
```{r}
blbees <- read.csv(path.expand("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/BIO-Logic/table14_diversityevennessbees.csv"))

head(blbees)

ggplot(blbees, aes(Dist, Simpson))+
  geom_boxplot()


ggplot(blbees, aes(Dist, Evenness))+
  geom_boxplot()


ggplot(blbees, aes(Dist, PBO))+
  geom_boxplot()
```


