---
title: "Survey123_to_mySQL"
author: "Michelle DePrenger-Levin"
date: "2024-06-12"
output: html_document
---

```{r}
rm(list=ls())

library(ggplot2)
library(tidyr)
library(dplyr)
library(stringr)

library(tidyverse)
library(sf)
library(mapview)
library(leaflet)

# library(remotes)
# remotes::install_github("r-spatial/mapview@429-update-to-webshot2")
```



2024 was the first year we collected data using a Survey123 form. The data needs to be reformatted to ingest by the PHP mySQL database.   
```{r}

currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))



## Transect and Site, 
demo2024 <- read.csv("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/Sclerocactus_glaucus_DemographicMonitoring_0.csv")

plants2024 <- read.csv("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/plantData_1.csv") 

sitesTransID <- read.csv("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/SitesTransectsDate.csv")

## Column names for annual data direct entry into dbg-research-db  
scgl_data_columns <- read.csv("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/_scgl_data.csv")
scgl_tags_columns <- read.csv("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/_scgl_tags.csv")

```






## Needed edits 
ScGl_tag_id  Site    transect    Site
528          5	     19	         Atwell Gulch - 108      should be 377 as Tag and the rest zero, add the tag_id     
   
More edits below once connect with known X and Y coords   
```{r}
plants2024$ScGl_tag_id[plants2024$Height == 377 & !is.na(plants2024$Height)] <- "528"
plants2024$Tag[plants2024$Height == 377 & !is.na(plants2024$Height)] <- "377"
plants2024$Height[plants2024$Tag == 377 & !is.na(plants2024$Height)] <- 0

plants2024 %>%
  filter(ScGl_tag_id == 528)


plants2024$Tag[plants2024$ObjectID == 3] <- 185.07

## A second measurement of the same plant
plants2024 <- plants2024 %>%
  filter(ObjectID != 1183)

## All NAs
plants2024 <- plants2024 %>%
  filter(ObjectID != 147)

## Plant 280.2 assigned to the wrong tag, should be new, not existing 
plants2024 <- plants2024 %>%
  mutate(Tag = case_when(ObjectID == 69 ~ "280.02",
                         TRUE ~ Tag)) 



```

# More edits found after merging
objectID 147 needs to be deleted  - it is gone (2024-07-17)
ObjectID 75 is the same plant recorded elsewhere  --> should be recorded under 180   
   
   180 = X 0.81, Y 12.5      
   
     180: tag 12 cm SW (TagID 437) Years: 2009:2016   
     180.01: tag 5cm W (TagID 3641) Years: 2021-2022   
     180.02: tag 9 cm SW (TagID 3642) Years: 2021-2022    
     ObjectID 69 should be assigned to tag 180.03 NEW
     
   280 = X 0.9, Y 12.86 (ObjectID 75)    
   
     280.02: tag 30cm S (TagID 467) Years: 2011:2023     
     
```{r}
plants2024 %>%
  filter(ObjectID %in% c(147,75,69))

## If plant 280.2 should be same as 180, figure out which 180 and remove this tag from the database
plants2024 %>%
  full_join(demo2024, by = c("ParentGlobalID" = "GlobalID")) %>%
  full_join(sitesTransID, by = c("transect" = "TransectID")) %>%
  distinct() %>%
  filter(ObjectID.x %in% c(69,75))



### Tag 280.2 must mean 280.02 with Tag 30cm S is the same as 180
plants2024 <- plants2024 %>%
  filter(ObjectID != 75)

## However, ObjectID 69: 280.02 tag 10cm SW is too small to be old 280.02, Could be a new one, remove ScGl_tag_id, make it 280.09
plants2024 <- plants2024 %>%
  mutate(ScGl_tag_id = case_when(ObjectID == 69 ~ "NA",
                                 TRUE ~ ScGl_tag_id)) %>%
  mutate(Tag = case_when(ObjectID == 69 ~ "280.09",
                                 TRUE ~ Tag))
  
```
#########################################################

# Select Option 2 until the upload can be altered to match the output from Survey123
# Option 1 - direct upload into the PHP MySQL database       

Remaining questions:    
      1. how to deal with the ScGl_data_id?     
      2. Note that "PhotoNumber" is now used to record the number of flowers or fruits     
      3. How to update (instead of add on) any changes to X, Y, or location, or tag number?   
      4. What happens to the old_tag_id... through duplicate_tag_id_2?
          Can update column by column so only need affected ones
      5. Save an update file with old location, the proposed new location, and the comments. Do quality control on the proposed new location, then upload to update. 
      6. Need a way to check that the update does what we think it's doing. Check that it matches. 
      
      
      

______________________________________________________________________________________________________________
# Option 2 - format to the paper2csv DBGDatabse upload   
Work site by site, then can have NEW 

To upload to MySQL through the PHP through the web format       

Step 1   
   download current year csv from database, save as ToEnter_2024     
Step 2   
   push to git
Step 3
   Use VLOOKUP in excel to put data2pasteincsv_2024.csv into the correct lines
   Save the "NEW" lines by hand
   Update the fixXYbyhand_2024.csv and NEWdecimalXY_2024.csv by hand
   Hopefully VLOOKUP can change only where matches for IndivPlantLocation2Update_2024.csv
   
Step 3a
     "IndivPlantLocation2Update_2024_v2.xlsx" has X and Y corrected and NewIndivdLocation with appropriate information from concatenated IndivdualPlantLocation and Comments.     
     
     ".\Denver Botanic Gardens\Conservation - General\AllProjectsBySpecies\Sclerocactus-glaucus\Sclerocactus-glaucus_DataEntry\2024_Sclerocactus-glaucus_Dataentry_Survey123-to-php\IndivPlantLocation2Update_2024_v3_vlookup.xlsx" is a copy with correct X and Y and will be used to VLOOKUP into the final document to upload to MySQL database.      
      1. Bridgeport TagID 3925 row has been deleted because it should not appear in transect 3, is correctly in transect 2  


   
```{r}
## Add year to each data column 

## Needs to skip 3 non-blank rows to get to headers
headers <- read.csv(paste0("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/ToEnter_", currentyr, ".csv"), 
                    skip = 3,
                    header = FALSE, nrows = 1, as.is = TRUE)
# Only skip 4, it will ignore the top 3 blank rows
df2enter <- read.csv(paste0("https://raw.githubusercontent.com/DenverBotanicGardens/ScGl/master/datasets/ToEnter_", currentyr, ".csv"),, skip = 4, header = FALSE)

# Match gives first time it matches
headers[,match("Height (cm)", headers)] <- paste0("Height (cm)", currentyr - 2)
headers[,match("Width (cm)", headers)] <- paste0("Width (cm)", currentyr - 2)

headers[,match("Height (cm)", headers)] <- paste0("Height (cm)", currentyr - 1)
headers[,match("Width (cm)", headers)] <- paste0("Width (cm)", currentyr - 1)

headers[,"Comments" == headers][2] <- paste0("Comments", currentyr)

## remove all white space and special characters
headersNowhite <- sapply(headers, function(x) gsub(" ", "", x))
headersNospch <- sapply(headersNowhite, function(x) str_replace_all(x, "[^[:alnum:]]", ""))

colnames(df2enter) <- headersNospch

# Data frame to enter current year data into
## Use NA to advantage to remove rows without Transect info
# Need to keep IndividualPlantLocation to update the existing location information with appropriate parts from current comments
df2enter <- df2enter %>%
  select(c(TagIdignoreforupload:Ycoordm, IndividualPlantLocation)) %>%
  mutate(Transect = as.numeric(as.character(Transect))) %>%
  mutate(MySQLTransectID = as.numeric(as.character(MySQLTransectID))) %>%
  filter(!is.na(Transect) & Transect > 0) %>%
  mutate(TagIdignoreforupload = as.numeric(as.character(TagIdignoreforupload   )))
  # mutate(across(TagIdignoreforupload:Ycoordm, as.numeric)) %>%
  ## need to remove the NEW rows
  filter(!is.na(TagIdignoreforupload))

str(df2enter)
head(df2enter)

## Some have X and Y coordinates but that don't make sense. X_coord = 282 or 257
df2enter %>%
  filter(is.na(Xcoordm)) 

df2enter %>%
  filter(Xcoordm > 1)

```


Keep all rows from df2enter to pull the X Y even when the base tag doesn't have data this year. 
Check that all columns Height:Comments are filled out for the current year.
Export only rows that have data from this year (either a Height or a number of Minis)   



```{r}

## There must be 251 new plants or these are in error
table(plants2024$ScGl_tag_id[is.na(plants2024$ScGl_tag_id)], useNA = "always")

sitesTransID %>%
  filter(TransectID %in% c(8,119,121))


####################################################################################### Done in 2024_scGl_ToIngest_v2.csv
## What was going on with 142/442. It is recorded as tag number 142, should be changed to 442
# Need to do in the upload csv 
plants2024 %>%
  filter(Tag == "142/442")
df2enter %>%
  filter(Tag == "142")

plants2024 %>%
  filter(is.na(ScGl_tag_id) & is.na(Height)|is.na(ScGl_tag_id) & is.na(X..Minis)) 

## Probably 189 new plants
plants2024 %>%
  filter(is.na(ScGl_tag_id) & !is.na(Height)|is.na(ScGl_tag_id) & !is.na(X..Minis))%>%
  summarise(NumNew = n())


##### change data for tag_id 3543 to 3366 so data goes in there!!!! ############
plants2024 %>%
  filter(ScGl_tag_id %in% c(3543,3366))
plants2024 <- plants2024 %>%
  filter(!ScGl_tag_id == 3543)


# Quality control, some errors in data entry
## I assume if there was no ScGl_tag_id that the Tag is new  
## These don't have X and Y of the tags. Need the data upload csv from the DBG Database for those; df2enter has info from database, if new, will not have Tag.y and will not have   
## Some duplicates, why? 
annualdata2QC <- plants2024 %>%
  select(Tag:ParentGlobalID) %>%
  # mutate(ScGl_tag_id = as.numeric(as.character(ScGl_tag_id))) %>% ## then numeric here and 
  left_join(demo2024, by = c("ParentGlobalID" = "GlobalID")) %>%
  left_join(sitesTransID, by = c("transect" = "TransectID")) %>%
  full_join(df2enter, by = c("ScGl_tag_id"="TagIdignoreforupload", "transect"= "MySQLTransectID")) %>%
  select(c(Tag.x,Tag.y,Height:ScGl_tag_id, site:transect, Site:IndividualPlantLocation)) %>%
  arrange(Site,transect,Tag.x)

nrow(annualdata2QC[duplicated(annualdata2QC),]) +
nrow(annualdata2QC[!duplicated(annualdata2QC),])
nrow(annualdata2QC)

annualdata2QC %>%
  summarise(n = sum(duplicated(.)))
annualdata2QC %>%
  summarise(n = sum(!duplicated(.)))
annualdata2QC <- annualdata2QC %>%
  filter(!duplicated(.))

annualdata2QC %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/annualdata2qualitycheck_", currentyr, ".csv", sep=""))


## See ".\Denver Botanic Gardens\Conservation - General\AllProjectsBySpecies\Sclerocactus-glaucus\Sclerocactus-glaucus_DataEntry\2024_Sclerocactus-glaucus_Dataentry_Survey123-to-php\2024_ScGl_vlookup_Survey123-MySQL_TESTINGVLOOKUP.csv" for details of how to do the vlookup

### Data to vlookup enter into the uploadable csv
## Make fixes to this! Then save and add to the document
## But maybe keep all the height with NA, we did record dead and spines
data2vlookup <- annualdata2QC %>%
  filter(!(is.na(Height) & Tag.x == "" & is.na(X..Minis) & is.na(Comments)))


## Should be no rows with no Tag but a height, But yes! We do want to add the comments where we saw a dead one. There are multiple duplicates, Why? 
data2vlookup %>%
  filter(is.na(Height) & !is.na(Comments)) %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/EditsNotesfromComments", currentyr, ".csv", sep=""))

  

## Tag.x are tag numbers entered in Survey123 in 2024, Tag.y are from the database

```



# 62 are Either NEW or errors
With data2QAQC I have a few duplicates!    
There are 254 without duplicates    
There are 198 without duplicates and only having tag information   
There are 187 that have height info but no ID (new??)   

Some with X and Y coordinates must be new tags. 
```{r}
data2QAQC %>%
  filter(is.na(ScGl_tag_id)) %>%
  # filter(Tag != "" & !is.na(Tag)) %>%
  filter(Height > 0) %>%
  distinct(ObjectID.x, .keep_all = TRUE) 
# %>%
#   summarize(n= n())

## 15 new ones? Or 15 needing x and y
data2QAQC %>%
  filter(is.na(ScGl_tag_id)) %>%
  filter(Height > 0) %>%
  distinct(ObjectID.x, .keep_all = TRUE) %>%
  # filter(grepl("Y ", Comments))
  filter(grepl(paste(c("Y ", "y ", "X ", "x "), collapse = "|"),
              Comments))


## Need to pull these out to replace the X and Y data


```


Notes:
ObjectID.x == 312; tag 762 is at Pond (id = 7); says X 23, Y 4.63. X is wrong. Should it be 0.23? 
    1171 X 0.44, Y 5.19. 


## Match formatting for the annual data  
## Setting to NULL acts to delete the column 
```{r}
toupload <- plants2024 %>%
  mutate(year = currentyr) %>%
  mutate(old_tag_id = NA) %>% 
  mutate(old_tag_id_2 = NA,
         old_pond_tag_id = NA,
         old_pond_transect_id = NA,
         duplicate_tag_id = NA,
         duplicate_tag_id_2 = NA) %>%
  select(c(ScGl_tag_id, year, Height, Width, Flowering, Browsing, X..Minis, Photo.Number, Comments, 
           old_tag_id:duplicate_tag_id_2))

## Make all NAs NULL - that's not so easy to do!
# toupload %>%
#   left_join()
#   write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/annualdata2upload_", currentyr, ".csv", sep=""))

## Need to remove any tag information out of comments and move to location in tags

```


Quality control for the Individual plant location information task in the spreadsheet and the X and Y. I'm going to make a data frame with the old location information, the as is comments, and the proposed location information. I will attempt to automate the extraction of tag information from the comments and X and Y updates from the comments. 

```{r}



```

######################################################### Work on formating for VLOOKUP


## make edits to columns (same as above) to cut and paste into upload csv

```{r}

data2QAQC <- plants2024 %>%
  mutate(Tag = as.numeric(Tag)) %>%
  left_join(demo2024, by = c("ParentGlobalID" = "GlobalID")) %>%
  left_join(sitesTransID, by = c("transect" = "TransectID")) %>%
  # left_join(df2enter, by = c("Tag" = "Tag",
  #                            "transect" = "MySQLTransectID")) %>%
# Keep all rows from df2enter so we have all tag info with X and Y
  full_join(df2enter, by = c("Tag" = "Tag",
                             "transect" = "MySQLTransectID")) %>%
  mutate(ScGl_tag_id = as.numeric(ScGl_tag_id)) %>%
  select(c(ObjectID.x,  Site, TagIdignoreforupload, Tag, IndividualPlantLocation, 
           transect,  Xcoordm, Ycoordm, Height:ScGl_tag_id, site:transect)) %>%
  distinct(Tag, Height, Width, Flowering, .keep_all = TRUE) ## To get rid of duplicates from no Tag and five "NEW"

## Did I run the 'Needed edits' ca. lines 83 on?  Should be 377 for Tag
plants2024 %>%
  filter(ScGl_tag_id == 528)

data2QAQC %>%
  arrange(transect, Tag ) 


data2QAQC %>%
  filter(Tag >= 185 & Tag < 186 & transect == 19)

## Probably the "NEW" rows?? There are five "NEW" at the end of each site, with no tag to merge with, creates duplicates
data2QAQC %>%
  filter(is.na(Tag) & !is.na(Height)) 
## Only one?
data2QAQC %>%
  filter(is.na(Tag) & is.na(Height)) %>%
  summarise(numNew= n())

## 15 needed new x and y from data2QAQC  # Now only one? TagID 2474
data2QAQC %>%
  filter(is.na(ScGl_tag_id)) %>%
  filter(Height > 0) %>%
  distinct(ObjectID.x, .keep_all = TRUE) %>%
  filter(grepl(paste(c("Y ", "y ", "X ", "x "), collapse = "|"),
              Comments.x))

## plants2024$ScGl_tag_id  missing from some, don't know why; 21 like this; 20 after deleting ObjectID 1183, a duplicate measurement
# A tagID in the database but that didn't connect in the Survey123   
## These all look good
data2QAQC %>%
  filter(is.na(ScGl_tag_id) & !is.na(TagIdignoreforupload)) %>%
  arrange(TagIdignoreforupload)

```

Tags that have a tagID in the database but no matching Scgl_tag_id in Survey123  
3723 is in twice with different Height and widths, maybe accidentally measured it twice: keep first ObjectID 1164 and delete 1183    


Assign ScGl_tag_id when it occurs in TagIdignoreforupload      
```{r}
## Tag number not recorded for one plant in transect 77 at Bridgeport 
data2QAQC <- data2QAQC %>%
  mutate(Ycoordm = case_when(ObjectID.x == 844 ~ ((30.37 + 42.20)/2),
                             TRUE ~ Ycoordm))
data2QAQC <- data2QAQC %>%
  mutate(Xcoordm = case_when(ObjectID.x == 844 ~ 0.5,
                             TRUE ~ Xcoordm))

data2QAQC <- data2QAQC %>%
  mutate(Comments.x = case_when(ObjectID.x == 844 ~ "Tag and XY not recorded, between 30.37 and 42.20, Cactus on West side of rock.",
                             TRUE ~ Comments.x))

## No data recorded for ObjectID 158
data2QAQC <- data2QAQC %>%
  filter(ObjectID.x != 158)

## Assign ScGl_tag_id 
dataWQAQC <- data2QAQC %>%
  mutate(ScGl_tag_id = case_when( (is.na(ScGl_tag_id) & !is.na(TagIdignoreforupload)) ~ TagIdignoreforupload,
                                  TRUE ~ ScGl_tag_id))

## New seedlings or new to us measurements
dataWQAQC %>%
  filter(is.na(ScGl_tag_id))

dataWQAQC %>%
  filter(is.na(ScGl_tag_id)) %>%
  arrange(Site, transect, Tag) %>%
  write.csv(paste("C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_DataEntry/2024_Sclerocactus-glaucus_Dataentry_Survey123-to-php/NEW_", currentyr, ".csv", sep=""))

dataWQAQC_v2 <- dataWQAQC %>%
  filter(ObjectID.x != 158) %>%
  mutate(TagIdignoreforupload = as.character(TagIdignoreforupload)) %>%
  mutate(TagIdignoreforupload = case_when(is.na(ScGl_tag_id) ~ "NEW",
                                          TRUE ~ TagIdignoreforupload))

dataWQAQC_v2 %>%
  filter(TagIdignoreforupload == "NEW")

```



## Check that all additional off a tag have same X and Y    
```{r}

dataWQAQC_v2 %>%
  filter(ObjectID.x == 69)

data4csv <- dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  distinct(BaseTag, transect, site, Xcoordm, Ycoordm, .keep_all = TRUE) %>%
  group_by(BaseTag, transect, site) %>%
  filter(n() > 1)

## Use VLOOKUP to replace all NEW with the correct X and Y  
dataFIXxy <- data4csv %>%
  group_by(BaseTag, site, transect) %>%
  mutate(XcoordmNEW = max(Xcoordm, na.rm = TRUE)) %>%
  mutate(YcoordmNEW = max(Ycoordm, na.rm = TRUE))

dataFIXxy %>%
  filter(Xcoordm != XcoordmNEW) %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/ErrorsXY_", currentyr, ".csv", sep=""))

dataFIXxy %>%
  filter(TagIdignoreforupload == "NEW") %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/NEWdecimalXY_", currentyr, ".csv", sep=""))

## 6 errors, two have living plants, ObjectID 983 says on trans 3, should be 2; objectID 89 is dead spines; 
# ObjectIDs: 85, 89, 91, 93, 94, 983  
# Only BaseTag 263  and 238 have XY; some of 238 is in transect 21, recorded in 26 Survey123; 263 has in trans 77, new plants recorded in trans 76  ** NEED TO FIX IN DATABASE **
dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  distinct(BaseTag, transect, site, Xcoordm, Ycoordm, .keep_all = TRUE) %>%
  group_by(BaseTag, transect, site) %>%
  mutate(Xcoordm = max(Xcoordm, na.rm = TRUE)) %>%
  mutate(Ycoordm = max(Ycoordm, na.rm = TRUE)) %>%
  ungroup() %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/data2pasteincsv_", currentyr, ".csv", sep=""))

## Error in  group 545: BaseTag = NA, transect = 77, site = 10. Tag 545 is in transect 142, site 
## One at Bridegport (transect 77, is a small one with no tag number recorded, no X or Y)   
dataWQAQC %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  filter(is.na(BaseTag) & !is.na(Height))

## Must be within range of surrounding records
# NA is Y between 30.37 and 42.20, 
dataWQAQC %>%
  filter(ObjectID.x > 800 & ObjectID.x < 900)

### Oh, to get X and Y, need to pull all tags from download, not just the ones with data this year. 
## Just NEW to paste in bottom. Otherwise only takes the first value and puts in every NEW slot. Need to cut and paste
dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  distinct(BaseTag, transect, site, Xcoordm, Ycoordm, .keep_all = TRUE) %>%
  group_by(BaseTag, transect, site) %>%
  mutate(Xcoordm = max(Xcoordm, na.rm = TRUE)) %>%
  mutate(Ycoordm = max(Ycoordm, na.rm = TRUE)) %>%
  filter(TagIdignoreforupload == "NEW") %>%
  ungroup() %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/NEW2pasteincsv_", currentyr, ".csv", sep=""))

## Didn't pull X Y for these:
# Missing 137, no X and Y. Only 137.01 is in the export above. Where is 137.02???
dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  filter(BaseTag == 137)

## Only the NEW 166.03 and 166.04
dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  filter(BaseTag == 166 &
           transect == 23)


## Check errors
dataWQAQC_v2 %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  filter(BaseTag %in% c(266,241,238,279,443,263))

## Need to pull X and Y from comments for NEW tags
## Fix by hand
dataWQAQC_v2 %>%
  filter(grepl(paste(c("X "," X", "x ", " x"), collapse = "|"),
              Comments.x)) %>%
  write.csv( file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/fixXYbyhand_", currentyr, ".csv", sep=""))
  

## Are there any that are not new but don't match? 
dataXYerrors <- dataWQAQC_v2 %>%
  filter(TagIdignoreforupload != "NEW") %>%
  mutate(BaseTag = trunc(Tag, prec = 0)) %>%
  distinct(BaseTag, transect, site, Xcoordm, Ycoordm, .keep_all = TRUE) %>%
  group_by(BaseTag, transect, site) %>%
  filter(n() > 1)

dataFIXxy_errors <- dataXYerrors %>%
  group_by(BaseTag) %>%
  mutate(Xcoordm = max(Xcoordm, na.rm = TRUE)) %>%
  mutate(Ycoordm = max(Ycoordm, na.rm = TRUE))

```



# Will update myPHPAdmin to match data collected in the field from Survey123
#### 19 plants that have a ScGl_tag_id from plants2024 in the Survey123 but no similar record in the phpMyAdmin database   


454 was big, saw dead spines, 445 not in transect 21 at Atwell Gulch, IS in transect 20   
3829 need to update X and Y   
3903 is in transect 21 in Survey123 but 20 in database   
3905 ""   
442 ""    
3639 ""   
473 ""   
443 ""   
472, 493, 474, 444, 2583, 2961. 3266 ""   
3925 is in 76 in the Survey123 but in 77 in the database.   
3926, 3927, 3930    
3927 is repeated for tag 263.03 and 263.02; is with 263.02 in database. maybe 263.03 is NEW? No, was just under 77
3930 is associated with tag number 263.05 in the database     
    3925: 263   
    3926: 263.01   
    3927: 263.02   
    3928: 263.03   
    3929: 263.04    
    3930: 263.05   


## Individual plant location to update
```{r}

# dataWQAQC_v2 %>%
#   filter(grepl(paste(c("tag ", "Tag "), collapse = "|"),
#               Comments.x)) %>%
#   write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/IndivPlantLocation2Update_", currentyr, ".csv", sep=""))


# ?? Pull out X and Y from location info and just general info? No way to do it systematically because we didn't collect the info in a consistent way

## Just tack on the comments if there is 'tag', assume we typed it in because it needed updating
dataWQAQC_v2 %>%
  filter(grepl(paste(c("tag ", "Tag "), collapse = "|"),
              Comments.x)) %>% 
  unite(NewIndivdLocation, c("IndividualPlantLocation", "Comments.x"), sep = "; ", remove = FALSE) %>%
  write.csv(  file = paste("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/hackathon/ScGl/datasets/IndivPlantLocation2Update_", currentyr, ".csv", sep=""))
  
```


# NEW plants to cut and paste into input sheet  
```{r}

dataWQAQC_v2 %>% 
  

```



New:
368.12 - 368.13 (should be 368.11 and 368.12) transect 18
376.01-376.02 trans 18
439.03-439.06, trans 18
694.05-694.07, trans 18
332.03, trans 17


Duplicates:
ScGl_tag_id 2254 and 2476 for 628.01; 2476 no more data with it ** Deleted from MySQL database **

            





```{r}

m <- demo2024 %>%
  left_join(sitesTransID, by = c("transect" = "TransectID")) %>%
  mapview(  xcol = 'x', ycol = 'y', crs = 4269, grid = FALSE,
            zcol = 'Site')

mapshot(m, file = "C:/Users/deprengm/Denver Botanic Gardens/Conservation - General/AllProjectsBySpecies/Sclerocactus-glaucus/Sclerocactus-glaucus_AnnualReports/DemographicSitesMap.png")
```



Merge current year data
```{r}

demo2024 %>%
  select(c(GlobalID:transect, EditDate)) %>%
  left_join(plants2024, by = c("GlobalID" = "ParentGlobalID")) %>%
  select(c(Tag, transect, Height:Comments))   

## Need to move any tag information from "Comments" to the "Individual Plant Location" 
df2enter %>%
  unite(IndividPlLoc_comments, c("IndividualPlantLocation", "Comments"))
  # mutate(IndividualPlantLocation = case_when(grepl("tag", Comments) ~ 
  #                                                  paste(IndividualPlantLocation , Comments, sep = " New: "),
                                                 TRUE ~ IndividualPlantLocation ))


```

